---
title: "Class 16"
author: " "
header-includes:
   - \usepackage{pdfpages}
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Plan
* Announcements
* Review of Multivariate Regression
* Prediction and Forecasting (brief lecture)
* Forecasting Polls in R
* Intro to Group Assignment (Election Prediction)

# Review of Multivariate Regression

### Interpretation Practice

Your friend Cornelius wants to model how the density of wind farms influences support for environmental causes within a state. He also suspects that the wealthier the state, the higher the concern for environmental causes.

The variables are defined as follows:

- Support for Environmental Causes: Integer ranging from 0 to 100, where 100 is equivalent to full support
- State Median Income: In thousand dollars (i.e. 47 = $47000)
- Wind Farms: Per 100,000 Population (i.e. 3 = 3 wind farms per every 100k people)

You obtain coefficients from a linear regression model:

Support for Environmental Causes = 20 + 0.2(State Median Income) + 2.4(Wind Farms)

Interpret the slope coefficients


### Review: Omitted Variable Bias

```{r}
library(tidyverse)
library(lubridate)
denmark <- read.csv("./class_data/denmark.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/denmark.csv","denmark.csv")
#denmark<- read.csv("denmark.csv")
```

To further understand multivariate slope coefficients, let's look at 4 variables in the Danish crime dataset:

violentcrime_rate
pctforeign
social_head
unemp_rate

First, let's examine the correlations between them:

```{r}

denmark %>% select(violentcrime_rate,pctforeign,unemp_rate,social_head) %>% 
  cor(use="complete.obs")

```

Now let's regress pctforeign on violentcrime_rate, and sequentially add additional variables:

```{r}

result <- lm(violentcrime_rate ~ pctforeign,data=denmark)
summary(result)

#result <- lm(violentcrime_rate ~ pctforeign +,data=denmark)
summary(result)

#result <- lm(violentcrime_rate ~ pctforeign +,data=denmark)
summary(result)

```

If something is uncorrelated with either X or Y, adding it to the equation will not change the coefficients. This is true by definition if X is randomized. Let's return to one of the first experiments we analyzed, on resume call back rates by (randomized) race. Here, race was randomized, so it should be uncorrelated with 'sex'.

```{r}

resume <- read.csv("./class_data/resume.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/resume.csv","resume.csv")
#resume <- read.csv("resume.csv")


resume$white <- ifelse(resume$race=="white",1,0)
resume$female <- ifelse(resume$sex=="female",1,0)

cor(resume$white,resume$female,use="complete.obs")
```

```{r}
result <- lm(call ~white, data=resume)
summary(result)

result <- lm(call ~white + female, data=resume)
summary(result)
```



Determining the strongest predictors:

violentcrime_rate          Violent crime (scaled by population)
lowincome_rate             Poverty Rate (scaled by population)
unemp_rate                 Unemployment Rate (scaled by population)
pct1565                    % of Population Aged 15-65 (scaled by population)
pctforeign                 % Non-citizen population (scaled by population)
sd_share                   Social Democratic Share of council seats
pop                        Population
eEducationhead             Spending on education, per capita
eHousinghead               Spending on public housing, per capita
eSocialhead                Spending on social expenditures, per capita
eTrafficInfrastructure     Spending on infrastructure, per capita
eTotalhead                 Total spending, per capita
seatsDPP                   Council seats held by far-right danish people's party


```{r}

ds <- denmark %>% select(violentcrime_rate,lowincome_rate,unemp_rate,pct1565,pctforeign,sd_share,pop,eEducationhead,eHousinghead,
                         eSocialhead,eTrafficInfrastructure,eTotalhead,seatsDPP)

ds %>% cor()

#result <- lm(violentcrime_rate ~ pctforeign + eTotalhead + eSocialhead,data=denmark)
#summary(result)

# Machine learning (demonstration only, do not need to know code)
library(glmnet)
X <- as.matrix(ds %>% select(-violentcrime_rate))
y <- as.numeric(ds$violentcrime_rate)
fit <- cv.glmnet(X, y, family="gaussian", alpha=1)
coef(fit)


#result <- lm(violentcrime_rate ~ ,data=denmark)
#summary(result)

```


# Prediction and Forecasting (brief lecture)

# Prediction in R

We can predict outcomes in R using predict(). Let's predict the violentcrime_rate for a municipality with
a population that is 2% foreign born (.02), and an unemployment rate of 10% (.1)

```{r}
result <- lm(violentcrime_rate ~ pctforeign + unemp_rate,data=denmark)
#new <- data.frame(oldvariablename = newdata)
predict(result, newdata=new)
  
```

Imagine that it is the day before the election in 2012. We want to use polling data to predict the outcome of the _upcoming_ election.

In this situation, we would have to use historic data (2008) to forecast 2012 outcomes.

```{r}

polls08 <- read.csv("./class_data/polls08.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/polls08.csv","polls08.csv")
#polls08 <- read.csv("polls08.csv")

votes08 <- read.csv("./class_data/pres08.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/pres08.csv","pres08.csv")
#votes08 <- read.csv("pres08.csv")

votes08$actual.margin <- votes08$Obama - votes08$McCain
polls08$days.until <- ymd("2008-11-04") - mdy(polls08$middate)
polls08$margin <- polls08$Obama - polls08$McCain
```

*Step 1: Fit the training model*

A: Save the most recent poll within each state within a new dataframe called _polls08_latest_

```{r}

#polls08_latest <- polls08 %>%
```

B: Merge to the actual outcomes in 2008, and fit a bivariate model. Important: use the following lm() syntax: lm(y~x,data=dataframename)

```{r}

#merged08 <- polls08_latest  %>% left_join()
#result <- lm(,data=merged08)
summary(result)

```

*Step 2: Predict using test data*

Now, to predict outcomes in 2012, we need a similar set of independent variables from 2012:

```{r}
polls12 <- read.csv("./class_data/polls12.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/polls12.csv","polls12.csv")
#polls12 <- read.csv("polls12.csv")

polls12$margin <- polls12$Obama - polls12$Romney
polls12 <- polls12 %>% group_by(state) %>% arrange(days.until) %>% slice(1)
```

Create a new dataframe containing just the new independent variable(s). _We need give the variables the same name as in the original regression_

```{r}
# <- data.frame(#oldvariablename = newdata)
  
```

Now, let's predict outcomes in 2012 using the predict() function. Syntax: predict(#regression result vector, newdata = #newdataframe)

```{r}

```

We can relabel vectors using names():

```{r}
#names() <- 
```


# Intro to group assignment


## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?