---
title: "Class 18"
author: " "
header-includes:
   - \usepackage{pdfpages}
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Plan
* Quick discussion about group projects
* Probability & Uncertainty (Lecture)
* Central Limit Theorem Demonstration
* Standard error

# Basic probability theory (lecture)


# Probability practice (group work)

If the probability that the Democrats win State A is 47%, and the probability they win State B 
is 37%, what is the probability that they win both states?

```{r}

```

If the probability that the Democrats win State A is 47%, and the probability they win State B 
is 37%, what is the probability that they win either of the states?

```{r}

```

If the probability that the Democrats win State A is 47%, and the probability that the Green party wins State A 
is 2%, what is the probability that either the Democrats or the Green party win?

```{r}

```


# Uncertainty (lecture)

# Central Limit Theorem Demonstration

Biden Supporter     = 1
Non-Biden Supporter = 0 

The support for Biden in this particular state is 60%:

```{r}
state <- c(rep(1,60000),rep(0,40000))

hist(state)
```

Loop through each poll, one at a time, and obtain the mean

```{r}
# You do not need to understand this code
n <- 10000     # number of polls
ss <- 1000     # sample size of each poll
results <- rep(NA,n)
for (i in 1:n){
  poll <- sample(state,ss,replace=F)
  results[i] <- mean(poll)
}
```

Plot the means

```{r}
hist(results,xlim=c(.5,.7),col="lightgray",breaks=(n/2))
abline(v=mean(results),col="blue",lwd=3)
abline(v=.6,col="red",lwd=3)
```

We have n values: how many are outside the boundaries? 

If it's a perfectly normal distribution, it should be ~2.5% on each side.

```{r}
length(results[results < mean(results) - 1.96*sd(results)])  / n
length(results[results > mean(results) + 1.96*sd(results)])  / n
```


# How standard error works

Let's use the normal distribution to calculate the standard error of a poll of size 100.

First, let's simulate doing this manually:

```{r}
# You do not need to understand this code

population <- c(rep(1,60000),rep(0,40000))

# Simulation - Get sampling distribution
n <- 1000  # number of polls
ss <- 100  # sample size
results <- rep(NA,n)
for (i in 1:n){
  poll <- sample(population,ss,replace=F)
  results[i] <- mean(poll)
}

# Sampling distribution
hist(results,breaks=50)

# Standard deviation of sampling distribution
round(sd(results),2)
```

Now let's use the formula (with only 1 poll)

```{r}
poll <- sample(population,ss,replace=F)
p_sd <- sd(poll)
se <- p_sd / sqrt(ss)
round(se, 2)
```

# Group Work: Error in Average Treatment Effects

First, obtain the difference in means between the treated and control group for the _postattack_ variable.

Treated group: (fire==1)
Control group: (fire==0)

```{r}
library(tidyverse)

chechen <- read.csv("./class_data/chechen.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/chechen.csv","chechen.csv")
#chechen <- read.csv("chechen.csv")

```

Estimate the standard error for this difference in means, by obtaining the standard deviation and sample size of _each_ distribution. Remember that you can use table() to count; you can also use n() inside summarize(). To obtain the standard deviations for each group, use sd() on a filtered dataset, or use sd() within summarize()

```{r}
# Sample Size
n_treatment <- 
n_control <- 
  
# Obtain standard deviation

  
# Combine using formula
# sqrt( sd1^2 / n1     + sd2^2 / n2 )
  
# Answer should match: 0.4165658
```


# As a class (only if time):

For a _binary treatment indicator_, standard errors for the ATE can be quickly calculated using linear regression. Let's regress fire on postattack

```{r}
result <- lm()
summary(result)
```


## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?


