---
title: "Class 11"
author: "Your name"
header-includes:
   - \usepackage{pdfpages}
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Plan
* Midterm/Week's Schedule
* Short Lecture - Correlation
* Group Work on Correlation
* Short Lecture - Survey sampling
* Sampling error in R


# Short Lecture: Correlation

# Exploring descriptive relationships in data

Descriptive correlations can provide helpful information to policymakers (remember that any causal relationship between two variables will have a non-zero correlation). In this exercise, you will evaluate the factors that are correlated with violent crime, using data from Denmark.

```{r}
denmark <- read.csv("./class_data/denmark.csv")


#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/denmark.csv","denmark.csv")
#denmark<- read.csv("denmark.csv")

```

Variables:

violentcrime_rate          Violent crime (scaled by population)
murderrate                 Murders (scaled by population)
lowincome_rate             Poverty Rate (scaled by population)
unemp_rate                 Unemployment Rate (scaled by population)
pct1565                    % of Population Aged 15-65 (scaled by population)
pctforeign                 % Non-citizen population (scaled by population)
sd_share                   Social Democratic Share of council seats
pop                        Population
eEducationhead             Spending on education, per capita
eHousinghead               Spending on public housing, per capita
eSocialhead                Spending on social expenditures, per capita
eTrafficInfrastructure     Spending on infrastructure, per capita
eTotalhead                 Total spending, per capita
drug_total                 Total drug offenses
murder                     Total murders
violentcrime               Total violent crimes


Let's explore potential correlations between violent crime and other municipal-level variables using a scatter plot.

```{r}
y <- denmark$violentcrime_rate
x <- 
  
plot()
```

We can add a trend line to the plot using lines(lowess(x,y))

```{r}
```

We can also evaluate the correlation numerically using cor(x,y,use="complete.obs"):

```{r}
```

# Group work

With your neighbors, explore other correlations. Which factors are most strongly correlated with violent crime? Use a separate code chunks for each variable you explore. 

```{r}
```

```{r}
```

```{r}
```


# Short Lecture - Sampling Error


## Sampling and Survey Error

Supports Biden     = 1
Does not support Biden = 0 

The support for Biden in this particular state is 40%:

```{r}
state <- c(rep(1,40000),rep(0,60000))

```

To take a sample, we use the _sample()_ function. The format is:

  sample(whatwewanttosample,sizeofsample,replace=F)

Let's draw a sample of size 10, and examine the mean.

```{r}
sample(,,replace=F)

```

Now let's draw new samples of 100 and 500, and compare the mean.

```{r}

```

Now let's draw a 1000 person sample. What is the expected error, using the equation covered in class?

```{r}


```

To test whether our formula for the expected error works, let's draw 100 samples, each of size 1000. Then we'll plot the average value from each sample. A distribution of sample means is known as a *sampling distribution* 

```{r}
# You do not need to know this code
n<-1000
samples <- 100
out <- c()
for (i in 1:samples){
  out <- c(out,mean(sample(state,n,replace=FALSE)))
}

```

A polling company randomly sampled 500 Americans to ask them if they approved supplying arms to Ukraine. On a scale of 1 to 5, where 1 meant "strongly oppose" and 5 meant "strongly support", the average response was 4.12. The standard deviation was 0.25. What was the expected error of this poll?

```{r}

```

# Only if time: item non-response

Let's load a survey of Afghan citizens. 

```{r}
library(tidyverse)
afghan <- read.csv("./class_data/afghan.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/afghan.csv","afghan.csv")
#afghan<- read.csv("afghan.csv")

```

One of the most common items individuals refuse to respond to in surveys are questions about income. How many fall into this category in the afghan survey?

To determine this, we can use is.na() on the _income_ variable.

```{r}

```

How does non-response vary by other characteristics? Let's look at responses to the income question as a function of employment

```{r}

```


## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?

