---
title: "Class 10"
author: "Your name"
header-includes:
   - \usepackage{pdfpages}
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Plan
* DiD Review
* Implementing DiD in R
* Group Work - DiD

# Difference in Differences in R

```{r}
library(tidyverse)
minwage <- read.csv("./class_data/minwage.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/minwage.csv","minwage.csv")
#minwage <- read.csv("minwage.csv")

```

Let's create three new variables in the dataset:

a) The proportion of full-time workers before the reform
b) The proportion of full-time workers after the reform
c) The longitudinal difference between (a) and (b)
 

```{r}



```

To make the subsequent code simpler, let's create two filtered data frames, one for NJ and one for PA

```{r}

nj<-

pa <-
  
```

Did wages change in NJ? Let's compare the distribution of min wage, before and after the reform. 

```{r}


```

Difference in Differences Step 1: We need the average longitudinal difference of the outcome, for NJ and PA.

```{r}

```

Difference in Differences Step 2: Subtract the Control group from the Treated group

```{r}

```

Difference in Differences Step 3: Evaluate parallel trends

```{r}

```


# Another Example

In many (but not all) states in Austria, voting in parliamentary elections was compulsory, punishable by fines. In 1992, the Federal Constitutional Court announced that this practice was unconstitutional and struck down the law. 

We are interested in finding out whether the repeal of the law (and the accompanying reduction in turnout) reduced support for left-wing parties. States that were forced to abolish compulsory voting are considered the treatment group, while states that never enacted compulsory voting are in the control group. 

Import the dataset using the following commands:

```{r}
austria <- read.csv("./class_data/austria.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/austria.csv","austria.csv")
#austria <- read.csv("austria.csv")

```

The dataset contains a partial sample of election results from Austrian municipalities, and includes following variables:

name                    name of the municipality
state                   name of the state
treated                 1 if the state practiced compulsory voting prior to 1992, 0 otherwise
year                    election year
socialist_voteshare     the percentage of votes cast for the socialist party in an election year

This data is in a slightly different format than the minwage data; it is saved in long format


## Demonstration: Plotting

It is often helpful to plot the trends when doing DiD problems. Here is how we do so. First, create two new datasets that contain the average value of the outcome for each time period:

```{r}

treat.avg <- austria
  

control.avg <- austria

```

Next, use the plot(x=,y=,type="l") command to plot the time series.

```{r}

```

Let's adjust the vertical axis and add color:

```{r}

```


We can overlay the other data by using lines(x=,y=)

```{r}

```


# Group Work

1. Use the difference-in-differences estimator to evaluate the effect of repealing compulsory voting on municipal-level vote share for the Socialist Party. Interpret the results -- was the repeal of compulsory voting associated with reduced support for the Socialist party? 

 _Note: For this estimate, you should focus only on the elections held in 1990 (before) and 1994 (after), since the repeal occurred in 1992_ 
 
_There are several ways to proceed with the code. One of the simplest approaches is to calculate 4 averages: treat_post, treat_pre, control_post, and control_pre, and then calculate the DiD as: (treat_post - treat_pre) - (control_post - control_pre). I've started this code for you below_

```{r}
treat_post <- # obtain average of the outcome for the treated group in the post period

treat_pre <-
  
control_post <-
  
control_pre <-
  
```

2. To evaluate the parallel trends assumption, repeat the analysis from #1 using the years 1986 (before) and 1990 (after). The DiD estimate for the parallel trends test should be close to 0. Does the parallel trends assumption appear to hold in this case? What does that imply for causality?

```{r}


```

*short answer here*

_Stretch question_

3. Benchmark the size of the DiD using the standard deviation of the longitudinal differnece in the control group

```{r}

```


If time remaining, you can either work on your problem set or leave early. 


## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand (leave blank if nothing)?

