---
title: "Class 15"
author: " "
header-includes:
   - \usepackage{pdfpages}
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

# Plan
* Review
* Multivariate Regression (lecture)
* Interpreting multivariate regression coefficients

# Review: Bivariate regression coefficients

Your friend Hildegard wants to model the relationship between each state's Republican vote share in the 2022 Congressional election, and support for Jo Jorgenson (the Libertarian candidate) in the 2020 Presidential election. She downloads data from the internet that is formatted as follows:

Republican Vote Share '22 : Ranges from 0 to 100, where 100 is equivalent to 100% vote share
Libertarian Vote Share '20 : Ranges from 0 to 1, where 1 is equivalent to 100% vote share

She runs a regression on this data, and obtains the following coefficients: 
(Republican Vote Share '22) = 48 + 12(Libertarian Vote Share '20) 

Interpret the intercept and coefficient on (Libertarian Vote Share '20) in sentence form.

# Review: R2

```{r}
denmark <- read.csv("./class_data/denmark.csv")
afghan <- read.csv("./class_data/afghan.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/denmark.csv","denmark.csv")
#denmark<- read.csv("denmark.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/class_data/afghan.csv","afghan.csv")
#afghan<- read.csv("afghan.csv")

afghan <- afghan %>% mutate(nonresponse=ifelse(is.na(income),1,0))
```

Recall the regression where we examined the relationship between the low income rate and the violent crime rate:

```{r}
x <- denmark$lowincome_rate
y <- denmark$violentcrime_rate

result <- lm(y~x)

```

What does the R2 mean: ?

Where does it come from?

```{r}

ds <- denmark %>% select(lowincome_rate,violentcrime_rate)

# Prediction from regression
ds$prediction_regression <- predict(result)

# Guessing the mean
ds$prediction_mean <- mean(ds$violentcrime_rate,na.rm=T)

# Total Sum of Squares 
sst <- sum((ds$violentcrime_rate - ds$prediction_mean)^2)

# Explained Sum of Squares
sse <- sum((ds$prediction_regression - ds$prediction_mean)^2)

sse/sst

summary(result)


cor(ds$violentcrime_rate,ds$lowincome_rate,use="complete.obs")^2

```


# Lecture: Multivariate Regression


# Multivariate Regression in R

Let's return to the Afghan survey non-response example, where we regressed employed on non-response to the income question:

```{r}
  
result <- lm(,data=afghan)
summary(result)

```

Now let's implement a multivariate regression and regress age, educ.years, and employed on non-response:

```{r}

result <- lm(,data=afghan)
summary(result)

```

###Practice by interpreting these coefficients:


employed:

age:

educ.years:


Why does the coefficient for employed change? Let's look at correlations:

```{r}

afghan %>% select(employed,age,educ.years) %>% cor()

```

Now let's add a variable to the dataset that is uncorrelated with employed:

```{r}
# You do not need to know this code
afghan$really_likes_skittles <- rbinom(nrow(afghan),size=1,prob=.001)

```

Let's add it to the regression. Does the coefficient change? Why or why not?

```{r}

result <- lm(nonresponse ~ employed,data=afghan)
summary(result)

result <- lm(nonresponse ~ employed + really_likes_skittles,data=afghan)
summary(result)

```

# Group work

Let's go back to the Danish crime data. Using the candidate variables below, try to maximize R2 by fitting a regression with up to *3* independent variables. Once you find a regression with a high R2, interpret the R2 and each coefficient. If you are not sure what scale a variable is measured on, use hist(variable) to examine the distribution.

Variables:

violentcrime_rate          Violent crime (scaled by population)
lowincome_rate             Poverty Rate (scaled by population)
unemp_rate                 Unemployment Rate (scaled by population)
pct1565                    % of Population Aged 15-65 (scaled by population)
pctforeign                 % Non-citizen population (scaled by population)
sd_share                   Social Democratic Share of council seats
pop                        Population
eEducationhead             Spending on education, per capita
eHousinghead               Spending on public housing, per capita
eSocialhead                Spending on social expenditures, per capita
eTrafficInfrastructure     Spending on infrastructure, per capita
eTotalhead                 Total spending, per capita
seatsDPP                   Council seats held by far-right danish people's party



```{r}
#result <- lm(data=denmark,violentcrime_rate ~ )
summary(result)
```

[Sentences here]


## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?
