---
title: "Problem Set 1"
author: ""
output:
  pdf_document: default
---

```{r setup, echo=F}
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

_Note_: You are welcome to work on the problem set before Friday's class. However, note that by the end of Friday's class we will cover all of the material you will need to complete the problem set.


# Bias in Self-reported Turnout

Surveys are frequently used to measure political behavior such as voter turnout, but some researchers are concerned about the accuracy of self-reporting. In particular, they worry about possible social desirability bias. In this phenomenon, respondents lie on surveys out of a desire to provide the 'socially correct' answer. In this case, those who did not vote in an election may report that they did. 

To study this issue, we can look at data from the ANES. The ANES is a nation-wide survey that has been conducted for every election since 1948. The ANES conducts face-to-face interviews with a nationally representative sample of adults.  

The following variables are in the turnout.csv file:

| Variable    | Description                                             |
|-------------|---------------------------------------------------------|
| year        | election year                                           |
| ANES        | Estimated turnout rate, based on self-reporting         |
| VEP         | Voting Eligible Population (in thousands)               |
| VAP         | Voting Age Population (in thousands)                    |
| total       | total ballots cast for highest office (in thousands)    |

## Loading the data

If you can sync with git, run the chunk below without modifying anything. If you cannot, uncomment the two following lines.

```{r}
turnout <- read.csv("data/turnout.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-w24/main/Problem%20Sets/data/turnout.csv","turnout.csv")
#turnout <- read.csv("turnout.csv")

```

## Questions

### Question 1

1.1. How many observations (rows) and variables (columns) are there in the data set? Use R functions. 

```{r}

```

1.2. What is the range of years covered in this data set? Note: you can look up basic R functions online, or using this pdf: https://cran.r-project.org/doc/contrib/Short-refcard.pdf

```{r}

```


### Question 2

2.1. Using the *$* symbol to reference specific columns in the dataset, create a new variable called **VAP_turnout** that contains the turnout rate for each year. To do so, divide the total votes by the voting age population. 

```{r}

```

2.2. Using your new variable, what is the average turnout rate in the data? What is the median turnout rate? 

```{r}

```

### Question 3

Now we will examine social desirability bias.

3.1. Create a new variable named *anes_vap_diff* that computes the difference between the actual turnout rate (calculated in 2.1) and ANES estimates of the turnout rate for each year.

```{r}

```

3.2. How large is the difference on average? 

```{r}

```

3.3 More recent versions of the ANES have focused on interviewing eligible voters. To see if this eliminates the discrepancy we observe, recalculate the average difference between the reported turnout rate and the voting *eligible* population. You'll need to repeat steps 2.1, 3.1, and 3.2 using different variables. Please do so in a single code chunk below:

```{r}

```

3.4. Viewed together, what do these results from 3.2 and 3.3 suggest? 

[Answer in full sentences here]

3.5. Open up the full dataset. Does social desirability bias appear to be larger or smaller in Presidential elections (84,88,92, etc), or in Congressional elections (86,90, etc)? What might explain this?

[Answer in full sentences here]


Please Knit and submit using Canvas
